---
title: Understanding the patterns of global re-emergence of RSV following COVID19 pandemic
author: "By Deus Thindwa"
date: "01/12/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
</style>

##### Load required packages for entire project
```{r LoadPackages, results = 'hide', message = FALSE, warning = FALSE}

pacman::p_load(char = c("lubridate", "tidyverse", "dplyr", "here", "rio", 
                        "scales", "boot", "magrittr",  "mvtnorm", "zoo", 
                        "patchwork", "PropCIs", "reshape2","purrr", "minqa", 
                        "ggridges", "timetk", "ggbreak", "ggpubr", "gridExtra", 
                        "curl", "archive", "jsonlite", "here"))
                        
``` 

##### Run the file cache (in silence)
```{r RunFileCache, results = "hide", echo = FALSE, message = FALSE, warning = FALSE}

##FROM excessILI package, Marcus Russi
now <- lubridate::now()

retrievePath <- function(fname, basepath='.', goalDate=lubridate::now()) {
  
  # Construct the path to the folder where all copies of 'fname' should be 
  # stored
  fullpath <- file.path(basepath, fname)
  
  # Make sure that this path is a directory I.e., archive/fname.txt needs to 
  # be a folder, not regular file.
  if (!dir.exists(fullpath))
    stop(sprintf("Path %s doesn't exist, or isn't a directory", fullpath))
  
  # Be sure that 'goalDate' is a Date object
  assertthat::assert_that(any("POSIXct" %in% class(goalDate)))
  
  # Get all the files in this directory
  dirListing <- list.files(fullpath)
  
  # If there were no files in the directory, we can't retrieve them!
  if (length(dirListing) == 0)
    stop(sprintf("No files were found in dirctory %s", fullpath))
  
  # Get a list of modification times for each file in the directory
  fullPaths <- file.path(fullpath, dirListing)
  mtimes <- purrr::lift_dl(file.mtime)(fullPaths)
  
  # The interval elapsed between the modified time of each file, and the 
  # goalDate. Our goal is to get the file with the smallest delta.
  absDeltas <-
    purrr::map_dbl(mtimes, ~lubridate::int_length(. %--% goalDate)) %>%
    abs
  
  # Find the index, and return it
  idx <- which(absDeltas == min(absDeltas))[1]
  
  fullPaths[idx]
}

#' Retrieve an RDS modified nearest to a specific date
#' 
#' Retrieves an RDS file stored in \code{basepath/fname} that has
#' the closest modification date to \code{goalDate}. If no such file exists, an
#' error will be thrown.
#'
#' @inheritParams retrievePath
#'
#' @examples
#' storeRDS(mtcars, 'mtcars')
#' identical(mtcars, retrieveRDS('mtcars'))
#'
#' storeRDS(mtcars, 'mtcars', basepath='.')
#' mtcars2 <- mtcars
#' mtcars2[,'cyl'] <- 1
#' storeRDS(mtcars2, 'mtcars')
#' mtcars_retrieved <- retrieveRDS('mtcars', goalDate = lubridate::now())
#' identical(mtcars, mtcars_retrieved)
#'
#' @export
retrieveRDS <- function(fname, basepath='.', goalDate=Sys.time())
  retrievePath(fname, basepath, goalDate) %>% readRDS

#' Store an R object into the file cache
#' 
#' Given an R object, attempts to store it in the directory
#' \code{basepath/fname}. The name given to the file will be of the form
#' \code{DATE.rds}, where \code{DATE} is of the format
#' \code{\%Y_\%m_\%d_\%H_\%M}.  An error will be thrown if \code{basepath} does
#' not exist. However, if \code{basepath/fname} does not exist, an attempt will
#' be made to create it. The \code{DATE} is the current time. Intended to be
#' used with \code{\link{retrieveRDS}}. See \code{\link{mostRecentTimestamp}}
#' for an usage example.
#' 
#' @param fname The name of the directory in \code{basepath} where various
#'   revisions of the file are stored. I.e., \code{file.txt} should be a
#'   directory, with revisions of the true \code{file.txt} stored inside of
#'   it.
#'
#' @param obj An R object
#'
#' @param basepath A string. The path which stores \code{fname}. Default '.'
#'
#' @return A message announcing the path the object has been written to
#' 
#' @examples
#' saveRDS(mtcars, 'cars')
#' saveRDS(mtcars, 'cars')
#' # Now the filesystem has, in '.':
#' # ├── mtcars
#' # │   ├── 2020_04_09_16_40.rds
#' # │   ├── 2020_04_09_16_41.rds
#'
#' @export
storeRDS <- function(obj, fname, basepath='.') {
  
  if (!dir.exists(basepath))
    stop(sprintf("Basepath '%s' does not exist. Cannot write file.", basepath))
  
  fullPath <- file.path(basepath, fname)
  
  # Create the directory for 'fname' if it doesn't exist. Notify the user.
  if (!dir.exists(fullPath)) {
    message(sprintf("Creating directory %s", fullPath))
    success <- dir.create(fullPath, recursive = FALSE, showWarnings = TRUE)
    
    if (any(!success))
      stop(sprintf("Failed to create directory %s", fullPath))
  }
  
  name <- as.character.POSIXt(Sys.time(), format="%Y_%m_%d_%H_%M.rds")
  writepath <- file.path(basepath, fname, name)
  
  saveRDS(obj, writepath)
  
  message(sprintf("Wrote object to %s", writepath))
}

#' Identify the timestamp of the most recently modified file in the file cache
#' 
#' Returns the timestamp of the most recently modified file. If no such file
#' exists, or if the directory \code{basepath/fname} doesn't exist, returns NA.
#' 
#' @param fname The name of the directory in \code{basepath} where various
#'   revisions of the file are stored. I.e., \code{file.txt} should be a
#'   directory, with revisions of the true \code{file.txt} stored inside of
#'   it.
#'
#' @param basepath A string. The path which stores the \code{fname} directory.
#'   Default '.'
#'
#' @return A POSIXct object specifying the \code{mtime} of the most recently
#'   modified file in \code{basepath/fname}
#'
#' @examples
#' library(lubridate)
#'
#' saveRDS(mtcars, 'cars')
#' saveRDS(mtcars, 'cars')
#' 
#' # Some time elapses...
#' 
#' # Decide if the latest version of 'mtcars' is "too old"
#' if (mostRecentTimestamp('mtcars') %--% now() > hours(24)) {
#'   # Store a "new" version
#'   saveRDS(mtcars, 'cars')
#' } else {
#'   cached_mtcars <- retrieveRDS('mtcars')
#' }
#'
#' @export
mostRecentTimestamp <- function(fname, basepath='.') {
  
  # Construct the path to the folder where all copies of 'fname' should be 
  # stored
  fullpath <- file.path(basepath, fname)
  
  # Make sure that this path is a directory I.e., archive/fname.txt needs to 
  # be a folder, not regular file.
  if (!dir.exists(fullpath))
    return(NA)
  
  # Get all the files in this directory
  dirListing <- list.files(fullpath)
  
  # If there were no files in the directory, we can't retrieve them!
  if (length(dirListing) == 0)
    return(NA)
  
  # Get a list of modification times for each file in the directory
  fullPaths <- file.path(fullpath, dirListing)
  mtimes <- purrr::lift_dl(file.mtime)(fullPaths)
  
  max(mtimes)
}

```

##### Run If Expired
```{r RunIfExpired, results = 'hide', message = FALSE, warning = FALSE}

#From Dan Weinberger (archiving)

# Using ExcessILI's data archiving functions, returns the most recent copy of
# output obtained by running a function or formula \code{f}, unless this 
# copy doesn't exist or is older (by modification time) than \code{maxage}.
# In that case, \code{f} is run and the output is archived into the folder
# Data/'storeName' as an RDS file, using the function ExcessILI::storeRDS.
#
# @param storeName A string. The name of the folder to store output in
# @param f A function or formula taking no arguments. Formulas are coerced to
#   functions.
# @param maxage How old can any existing archived file be before \code{f} is 
#   called again?

#source('./R/fileCache.R')
runIfExpired <- function(storeName, f, maxage=hours(0)) {
  basepath <- here("data")
  mostRecent <- mostRecentTimestamp(storeName, basepath=basepath)
  f <- rlang::as_function(f)
  
  runAndArchive <- function() {
    data <- f()
    storeRDS(data, storeName, basepath)
    data
  }
  
  if (is.na(mostRecent)) 
    return(runAndArchive())
  if (mostRecent %--% now() < maxage)
    return(retrieveRDS(storeName, basepath))
  runAndArchive()
}

```

##### Load RSV datasets
```{r LoadRSVdata, results = "hide", message = FALSE, warning = FALSE}

#read the WHO RSV update file into R
rsv <- runIfExpired('who_rsv', maxage = 168, ~read.csv(curl("https://frontdoor-l4uikgap6gz3m.azurefd.net/FLUMART/VIW_FNT?$format=csv")))

#get the required variables
rsvds <-
  rsv %>%
  dplyr::filter(!is.na(RSV)) %>%
  dplyr::select(WHOREGION, FLUSEASON, HEMISPHERE, COUNTRY_AREA_TERRITORY, MMWR_WEEKSTARTDATE, MMWR_YEAR, MMWR_WEEK, ORIGIN_SOURCE, RSV) %>%
  dplyr::arrange(WHOREGION, COUNTRY_AREA_TERRITORY, MMWR_WEEKSTARTDATE) %>%
  dplyr::rename("region" = WHOREGION,
                "fluseas" = FLUSEASON,
                "hemi" = HEMISPHERE,
                "country" = COUNTRY_AREA_TERRITORY,
                "date"= MMWR_WEEKSTARTDATE,
                "yr" = MMWR_YEAR,
                "wk" = MMWR_WEEK,
                "sentin" = ORIGIN_SOURCE,
                "cases" = RSV) %>%
  dplyr::mutate(date = date(date),
                sentin = factor(sentin),
                hemi = factor(hemi),
                fluseas = factor(fluseas))

rsvds_clean <- distinct(rsvds, country, date, .keep_all = TRUE)

#save the dataset as CSV
#rsvds %>% write_csv(here("data", "RSVglobal.csv"))

```

##### Yearly plotting RSV cases for Africa/South East Asia/Middle East/Western pacific

```{r PlotRSV1, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(country %in% c("Cameroon", "Central African Republic", "Côte d'Ivoire", "Madagascar",
                               "Mozambique", "Uganda", "South Africa", 
                               "India",
                               "Australia", "Japan", "Mongolia", "Malaysia",
                               "Oman", "Qatar")) %>%
  ggplot(aes(x = date(date), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```

##### Yearly plotting RSV cases in Europe

```{r PlotRSV2, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(
    country %in% c("France", "Germany", "Netherlands", "Spain", "Portugal", "Iceland",
                   "Ireland", "Denmark", "Finland", "Sweden", "Malta", "United Kingdom, England", 
                   "United Kingdom, Northern Ireland", "United Kingdom, Scotland",
                   "Bulgaria", "Belarus", "Russian Federation", "Hungary", "Poland", "Slovakia")) %>%
  mutate(country = if_else(country == "United Kingdom, Northern Ireland", "N Ireland",
                           if_else(country == "United Kingdom, Scotland", "Scotland",
                                   if_else(country == "United Kingdom, England", "England",
                                           if_else(country == "Russian Federation","Russia", country))))) %>%
  ggplot(aes(x = date(date), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```

##### Yearly plotting RSV cases in Americas

```{r PlotRSV3, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(
    country %in% c("Argentina", "Belize", "Bolivia (Plurinational State of)", "Brazil", "Canada", "Colombia", "Costa Rica",
                   "Dominican Republic", "Ecuador", "El Salvador", "Guatemala", "Honduras", "Mexico",
                   "Nicaragua", "Panama", "Paraguay", "Peru", "Uruguay")) %>%
  mutate(country = if_else(country == "Bolivia (Plurinational State of)", "Bolivia",
                           if_else(country == "Dominican Republic", "Dominica", country))) %>%
  ggplot(aes(x = date(date), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```

##### Monthly plotting RSV cases for Africa/South East Asia/Middle East/Western pacific

```{r PlotRSV4, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(country %in% c("Cameroon", "Central African Republic", "Côte d'Ivoire", "Madagascar",
                               "Mozambique", "Uganda", "South Africa", 
                               "India",
                               "Australia", "Japan", "Mongolia", "Malaysia",
                               "Oman", "Qatar")) %>%
  ggplot(aes(x = factor(month(date)), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```

##### Monthly plotting RSV cases in Europe

```{r PlotRSV5, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(
    country %in% c("France", "Germany", "Netherlands", "Spain", "Portugal", "Iceland",
                   "Ireland", "Denmark", "Finland", "Sweden", "Malta", "United Kingdom, England", 
                   "United Kingdom, Northern Ireland", "United Kingdom, Scotland",
                   "Bulgaria", "Belarus", "Russian Federation", "Hungary", "Poland", "Slovakia")) %>%
  mutate(country = if_else(country == "United Kingdom, Northern Ireland", "N Ireland",
                           if_else(country == "United Kingdom, Scotland", "Scotland",
                                   if_else(country == "United Kingdom, England", "England",
                                           if_else(country == "Russian Federation","Russia", country))))) %>%
  ggplot(aes(x = factor(month(date)), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```

##### Monthly plotting RSV cases in Americas

```{r PlotRSV6, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

rsvds_clean %>%
  dplyr::filter(
    country %in% c("Argentina", "Belize", "Bolivia (Plurinational State of)", "Brazil", "Canada", "Colombia", "Costa Rica",
                   "Dominican Republic", "Ecuador", "El Salvador", "Guatemala", "Honduras", "Mexico",
                   "Nicaragua", "Panama", "Paraguay", "Peru", "Uruguay")) %>%
  mutate(country = if_else(country == "Bolivia (Plurinational State of)", "Bolivia",
                           if_else(country == "Dominican Republic", "Dominica", country))) %>%
  ggplot(aes(x = factor(month(date)), y = cases)) +
  geom_line() + 
  facet_wrap(. ~ country, scales = "free_y") +
  theme_bw(base_size = 10, base_family = "Lato") +
  labs(title = , x = "MMWR Date", y = "RSV cases")

```