---
title: Understanding the patterns of global re-emergence of RSV following COVID19 pandemic
author: "Deus Thindwa"
date: "01/12/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```

&nbsp;

### Load required packages for entire project

```{r LoadPackages, results = 'hide', message = FALSE, warning = FALSE}
library(here)
source(here("script", "1_mainRSV.R"))
                        
```

&nbsp;

### Run the archiving script to update the RSV dataset

```{r RunFileCache, results = "hide", message = FALSE, warning = FALSE}

source(here("script", "2_fileCache.R"))
source(here("script", "3_runIfExpired.R"))

```

&nbsp;

### Load RSV datasets

```{r LoadRSVdata, results = "hide", message = FALSE, warning = FALSE}

source(here("script", "4_loadRSVdata.R"))

```

&nbsp;

### Seasonal dynamics of RSV by region

```{r PlotRSV1, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6, fig.align = 'center'}

source(here("script", "5_seasRSV_region.R"))

```

&nbsp;

### Seasonal dynamics of RSV in African, South East Asia, Western Pacific and Middle East countries

```{r PlotRSV2, message = FALSE, warning = FALSE, fig.width = 14, fig.height = 7, fig.align = 'center'}

source(here("script", "6_seasRSV_asmw.R"))

```

&nbsp;

### Seasonal dynamics of RSV in European countries

```{r PlotRSV3, message = FALSE, warning = FALSE, fig.width = 14, fig.height = 7, fig.align = 'center'}

source(here("script", "7_seasRSV_europe.R"))

```

&nbsp;

### Seasonal dynamics of RSV in Americas countries

```{r PlotRSV4, message = FALSE, warning = FALSE, fig.width = 14, fig.height = 7, fig.align = 'center'}

source(here("script", "8_seasRSV_americas.R"))

```

